# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
setwd('C:/Users/Jangedoo/Documents/GitHub/RepData_PeerAssessment1')
data = read.csv(file='activity.csv')
```

## What is mean total number of steps taken per day?
To compute this we will use sqldf library. It allows us to execute SQL queries in dataframe.
```{r warning=FALSE, message=FALSE}
library(sqldf)
query = 'select date, sum(steps) as dailySteps from data group by date'
transformed = sqldf(query)
transformed$dailySteps = as.numeric(transformed$dailySteps)
hist(x=transformed$dailySteps, xlab='Daily Steps',main='Histogram')

print(mean(transformed$dailySteps, na.rm=TRUE))
print(median(transformed$dailySteps, na.rm=TRUE))
```

## What is the average daily activity pattern?
To compute this, we just group-by the data on interval
```{r}
pattern = sqldf('select interval, avg(steps) avgSteps from data group by interval')
plot(pattern, type='l', xlab='Interval', ylab='Average no. of steps', main='Average number of steps per 5 minute interval')
```

The 5-minute interval with max. number of steps (on average). For this we just find out the interval with max number of avg. steps
```{r}
pattern[pattern$avgSteps == max(pattern$avgSteps), 'interval']
```

## Imputing missing values

Total Number of missing values in dataset.
```{r}
sum(!complete.cases(data))
```

To fill the missing values, we replace by the average for that 5-minute interval
```{r}
imputedData <- data
for(i in 1:nrow(imputedData)){
    if(is.na(imputedData[i, 'steps'])){
        imputedData[i, 'steps'] <- subset(pattern, interval == imputedData[i, 'interval'])$avgSteps[1]
    }
}
```

Histogram plot of imputed data
```{r}
query <- 'select date, sum(steps) as dailySteps from imputedData group by date'
transformed <- sqldf(query)
transformed$dailySteps <- as.numeric(transformed$dailySteps)
hist(x=transformed$dailySteps, xlab='Daily Steps',main='Histogram')

print(mean(transformed$dailySteps, na.rm=TRUE))
print(median(transformed$dailySteps, na.rm=TRUE))
```
## Are there differences in activity patterns between weekdays and weekends?
```{r}
weekend <- weekdays(as.Date(imputedData$date)) %in% c('Saturday', 'Sunday')
imputedData$dayType <- 'weekday'
imputedData$dayType[weekend == TRUE] <- 'weekend'
imputedData$dayType <- as.factor(imputedData$dayType)

weekdays <- subset(imputedData, dayType == 'weekday')
weekends <- subset(imputedData, dayType == 'weekend')

weekdays <- sqldf('select interval, avg(steps) as avgSteps from weekdays group by interval')
weekends <- sqldf('select interval, avg(steps) as avgSteps from weekends group by interval')
                  
par(mfrow=c(2,1))
par(mar = rep(2, 4))
plot(weekdays$interval, weekdays$avgSteps, type='l', main='weekday', xlab='interval', ylab='steps')
plot(weekends$interval, weekends$avgSteps, type='l', main='weekend', xlab='interval', ylab='steps')
```
